---
title: "Time series Analysis"
output: rmarkdown::github_document
---

Let us take a particular combination and analyze its sales.
For this purpose we first consider sales in store 1 of item 1.

```{r}
#Time series analysis
setwd("/media/aviator/Everything/Projects/Demand Analysis/Dataset")
dem_data=read.csv("train.csv")

library(lubridate)
dem_data$date=ymd(dem_data$date)

library(dplyr)
s1.i1=ts(filter(dem_data,(item==1 & store==1))$sales)
plot(s1.i1,main="sales of store 1 item 1",ylab="sales",col='blue',lw=3)


```

We  can see that the sale value ranges from about 2 to about 50. There is a very slow upward trend with time. There is a definite seasonality.
Let us now apply logarithmic transformation on the data to reduce its variance.

```{r}
#log transforming the time series to reduce variance
logd=log(s1.i1)  #without this step, values at ranged from -30 to 20, but now it ranges from 1.5 to 4.
plot(logd,main="log transformed data",ylab = "log of sales",col='blue',lw=2)
acf(logd,lag.max=50,main="ACF of log transformed data")
pacf(logd,lag.max=80,main="PACF of log transformed data")

```

After log transformation the data ranges from about 1.5 to 4. By looking at the ACF and PACF of the log data we can clearly see a seasonality at lag 7; i.e. seasonality of a period of week. PACF at lag 7 is much greater than seasonality at lag 1 hence we are not required to do non-seasonal differencing but only seasonal differencing at lag 7.

```{r}
final=diff(logd,7)
plot(final,main="log data, diff at intv 7",ylab = "log sales",col='blue')

```
The plot shows no visible trend. There are spikes in sales at different time points but they are more likely to be outliers than an indication of high variance.

```{r}
final=na.omit(final)
t1=decompose(ts(final,frequency = 7))
plot(t1$trend,main ="trend of final data" )
```
Decomposing the time series and looking at its trend also shows no visibly traceable trend in the data.

Now we shall look at the ACF and PACF plots to ascertain the parameters of our SARIMA model.
```{r}
acf(final,lag.max=50,main="ACF of seasonally differenced log data at interval 7")
pacf(final,lag.max=80,main="PACF of seasonally differenced log data at interval 7")

```

Since the ACF is slightly significant at lag 1 we may consider a non-seasonal moving average value of 0 or 1. Our seasonal moving average evidently value has to be 1 with period 7.
The PACF suggests a non-seasonal auto-regressive value of 1 since it is significant at lag 1 and then for about 9 periods of lag 7 we see spike in PACF suggesting a seasonal auto-regressive term ranging from 1 to 9.


Now let us perform Ljung-Box test to test stationarity of our final data.
```{r}
Box.test(final,log(length(final)))
```
The p-value is evidently much lower than alpha= 0.05 or alpha= 0.1 so the null hypothesis of 0 partial correlation coefficients (i.e. time series being stationary) is rejected.
Thus now we apply SARIMA model to our data.

```{r}
#trying different models
d=0
DD=1
aic=100000; ss=100000; order.1=NULL; order.2=NULL;
per=7

#value of p causing error. Look into variations for P.
#model failing whenever p>1.

for(p in 1:3){
  for(q in 1:3){
    for(i in 2:8){
      for(j in 2:3){
        if(p+d+q+i+DD+j<=19){
          model<-try(arima(x=logd, order = c((p-1),d,(q-1)), seasonal = list(order=c((i-1),DD,(j-1)), period=per)))
          if(class(model)=="try-error"){next()}
          pval<-Box.test(model$residuals, lag=log(length(model$residuals)))
          sse<-sum(model$residuals^2)
          cat(p-1,d,q-1,i-1,DD,j-1,per, 'AIC=', model$aic, ' SSE=',sse,' p-VALUE=', pval$p.value,'\n')
          if (model$aic<aic){
            aic=model$aic
            order.1=c((p-1),d,(q-1),(i-1),DD,(j-1))
          }
          if (sse<ss){
            ss=sse
            order.2=c((p-1),d,(q-1),(i-1),DD,(j-1))
          }
        }
      }
    }
  }
}


paste("Order with lowest aic is",list(order.1),"with aic value",aic)
paste("Order with lowest sse is",list(order.2),"with sse value",ss)

```
We see that to model the sales of item1 in store 1, the model with the lowest aic value of 234.6874 is SARIMA(1,0,1,1,1,1,7).
We see that to model the sales of item1 in store 1, the model with the lowest sse value of 117.7167743 is SARIMA(2,0,2,5,1,2,7).

